---
title: "NFM Parent doc"
author: "Kim Cressman"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(vegan))
library(glue)
library(goeveg)
```

Read and wrangle data  

```{r}
nfm_all <- read.csv(here::here("data", "NFM_counts_master_2020-06-09.csv"),
                     stringsAsFactors = FALSE)
```

Remove:  

+  rows where no sample was taken or no counts/lengths were recorded  
+  site 1, because exploratory analyses showed this site had very different salinity than the others  

Add:  

+  column of 1s, because NMDS doesn't work with rows of all 0s  
+  column to represent the year-season combination (to be used later in restricting permutations.)  

Mutate: turn the following columns into factors:  

+  collection_id
+  site  
+  habitat_type  
+  year_sampled  
+  season


```{r}
dat <- nfm_all %>% 
    filter(is.na(X00.no.sample.taken),
           is.na(X02.no.lengths..reason.unknown),
           site != 1) %>% 
    mutate(year_season = as.factor(paste(year_sampled, season, sep = "_")),
        across(c(collection_id, site, habitat_type, year_sampled, season), as.factor),
           ones = 1) %>% 
    relocate(year_season, .after = "season")
```

That took the data frame from 532 rows to 475.  


# Summary statistics  

Insert a child .Rmd here to generate all the summary info we want.  


# Habitat differences within each season: ANOSIM, SIMPER, and NMDS    

Setup - 

Performing transformations: square-root, and natural-log(x + 1). Adding 1 to keep negative values out of the matrices before generating Bray-Curtis matrices.  

```{r}
sqrt_dat <- dat %>% 
    mutate(across(Abudefduf.saxatilis:ones, ~sqrt(.x)),
           transformation = "sqrt")

log_dat <- dat %>% 
    mutate(across(Abudefduf.saxatilis:ones, ~log(.x + 1)),
           transformation = "ln(x+1)")
```

ANOSIM will be performed with restricted permutations, to only swap observations within a sample event (year-season combination). This lets us condition on the fish community available in the broader area at the time, and avoids issues like "what if spring was different in different years generally".  

In the code, this uses the `block` option. From the `permute` vignette:  

**Blocking factors which restrict permutations to within blocks.** This is what we need; allow shuffling only within the same sampling event. In this project, that's the combination of year and season.   

Also from the documentation: **Blocks are defined by a factor variable. Blocks restrict permutation of samples to within the levels of this factor; samples are never swapped between blocks.**  

***



## Spring  

```{r}
seas <- "Spring"
sqrt_sub <- filter(sqrt_dat, season == seas)
log_sub <- filter(log_dat, season == seas)
```

```{r, child = '02_ANOSIM-SIMPER-NMDS.Rmd'}

```

***  
***  


## Summer  

```{r}
seas <- "Summer"
sqrt_sub <- filter(sqrt_dat, season == seas)
log_sub <- filter(log_dat, season == seas)
```

```{r, child = '02_ANOSIM-SIMPER-NMDS.Rmd'}

```

***  
***  


## Fall  

```{r}
seas <- "Fall"
sqrt_sub <- filter(sqrt_dat, season == seas)
log_sub <- filter(log_dat, season == seas)
```

```{r, child = '02_ANOSIM-SIMPER-NMDS.Rmd'}

```

***  
***  


## Winter  

```{r}
seas <- "Winter"
sqrt_sub <- filter(sqrt_dat, season == seas)
log_sub <- filter(log_dat, season == seas)
```

```{r, child = '02_ANOSIM-SIMPER-NMDS.Rmd'}

```

***  
***  


