---
title: "NFM Parent doc"
author: "Kim Cressman"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(vegan))
```

# Read and wrangle data  

```{r}
nfm_all <- read.csv(here::here("data", "NFM_counts_master_2020-06-09.csv"),
                     stringsAsFactors = FALSE)
```

Remove:  

+  rows where no sample was taken or no counts/lengths were recorded  
+  site 1, because exploratory analyses showed this site had very different salinity than the others  

Add:  

+  column of 1s, because NMDS doesn't work with rows of all 0s  
+  column to represent the year-season combination (to be used later in restricting permutations.)  

Mutate: turn the following columns into factors:  

+  collection_id
+  site  
+  habitat_type  
+  year_sampled  
+  season


```{r}
dat <- nfm_all %>% 
    filter(is.na(X00.no.sample.taken),
           is.na(X02.no.lengths..reason.unknown),
           site != 1) %>% 
    mutate(year_season = as.factor(paste(year_sampled, season, sep = "_")),
        across(c(collection_id, site, habitat_type, year_sampled, season), as.factor),
           ones = 1) %>% 
    relocate(year_season, .after = "season")
```

That took the data frame from 532 rows to 475.  


# Summary statistics  

Insert a child .Rmd here to generate all the summary info we want.  


# Set up for ANOSIM, SIMPER, and NMDS  

Performing transformations: square-root, and natural-log(x + 1). Adding 1 to keep negative values out of the matrices before generating Bray-Curtis matrices.  

```{r}
sqrt_dat <- dat %>% 
    mutate(across(Abudefduf.saxatilis:ones, ~sqrt(.x)),
           transformation = "sqrt")

log_dat <- dat %>% 
    mutate(across(Abudefduf.saxatilis:ones, ~log(.x + 1)),
           transformation = "ln(x+0.1)")
```


***


# Habitat differences within each season  

## Spring  

```{r}
seas <- "Spring"
sqrt_sub <- filter(sqrt_dat, season == seas)
log_sub <- filter(log_dat, season == seas)
```

Insert child Rmd here for analyses  

```{r, child = '02_ANOSIM-SIMPER-NMDS.Rmd'}

```

***  
***  


## Summer  

```{r}
seas <- "Summer"
sqrt_sub <- filter(sqrt_dat, season == seas)
log_sub <- filter(log_dat, season == seas)
```

Insert child Rmd here for analyses  

```{r, child = '02_ANOSIM-SIMPER-NMDS.Rmd'}

```

***  
***  


## Fall  

```{r}
seas <- "Fall"
sqrt_sub <- filter(sqrt_dat, season == seas)
log_sub <- filter(log_dat, season == seas)
```

Insert child Rmd here for analyses  


```{r, child = '02_ANOSIM-SIMPER-NMDS.Rmd'}

```

***  
***  


## Winter  

```{r}
seas <- "Winter"
sqrt_sub <- filter(sqrt_dat, season == seas)
log_sub <- filter(log_dat, season == seas)
```

Insert child Rmd here for analyses  

```{r, child = '02_ANOSIM-SIMPER-NMDS.Rmd'}

```

***  
***  


