---
title: "NFM Parent doc"
author: "Kim Cressman"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(vegan))
suppressPackageStartupMessages(library(glue))
suppressPackageStartupMessages(library(goeveg))
suppressPackageStartupMessages(library(rgl))
suppressPackageStartupMessages(library(ggord))
suppressPackageStartupMessages(library(ggdist))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(patchwork))
suppressPackageStartupMessages(library(indicspecies))
suppressPackageStartupMessages(library(DT))
knitr::knit_hooks$set(webgl = hook_webgl)
```

Read and wrangle data  

```{r}
nfm_all <- read.csv(here::here("data", "NFM_counts_master_2020-06-09.csv"),
                     stringsAsFactors = FALSE)
```

Remove:  

+  rows where no sample was taken or no counts/lengths were recorded  
+  site 1, because exploratory analyses showed this site had very different salinity than the others  

Add:  

+  column of 1s, because NMDS doesn't work with rows of all 0s  
+  column to represent the year-season combination (to be used later in restricting permutations.)  

Mutate: turn the following columns into factors:  

+  collection_id
+  site  
+  habitat_type  
+  year_sampled  
+  season


```{r}
dat <- nfm_all %>% 
    filter(is.na(X00.no.sample.taken),
           is.na(X02.no.lengths..reason.unknown),
           site != 1) %>% 
    mutate(year_season = as.factor(paste(year_sampled, season, sep = "_")),
        across(c(collection_id, site, habitat_type, year_sampled, season), as.factor),
           ones = 1) %>% 
    relocate(year_season, .after = "season")
```

That took the data frame from 532 rows to 475.  


# Summary statistics  

```{r, child = '01_Summary_Stats.Rmd'}

```


# Habitat differences within each season   

Setup - 

Performing transformations: square-root, and natural-log(x + 1). Adding 1 to keep negative values out of the matrices before generating Bray-Curtis matrices.  

```{r}
sqrt_dat <- dat %>% 
    mutate(across(Abudefduf.saxatilis:ones, ~sqrt(.x)),
           transformation = "sqrt")

log_dat <- dat %>% 
    mutate(across(Abudefduf.saxatilis:ones, ~log(.x + 1)),
           transformation = "ln(x+1)")
```

ANOSIM will be performed with restricted permutations, to only swap observations within a sample event (year-season combination). This lets us condition on the fish community available in the broader area at the time, and avoids issues like "what if spring was different in different years generally".  

In the code, this uses the `block` option. From the `permute` vignette:  

**Blocking factors which restrict permutations to within blocks.** This is what we need; allow shuffling only within the same sampling event. In this project, that's the combination of year and season.   

Also from the documentation: **Blocks are defined by a factor variable. Blocks restrict permutation of samples to within the levels of this factor; samples are never swapped between blocks.**  

***



# Spring  

```{r}
seas <- "Spring"
sqrt_sub <- filter(sqrt_dat, season == seas)
log_sub <- filter(log_dat, season == seas)
```

```{r, child = '02_ANOSIM-SIMPER-NMDS.Rmd'}

```

***  
***  


# Summer  

```{r}
seas <- "Summer"
sqrt_sub <- filter(sqrt_dat, season == seas)
log_sub <- filter(log_dat, season == seas)
```

```{r, child = '02_ANOSIM-SIMPER-NMDS.Rmd'}

```

***  
***  


# Fall  

```{r}
seas <- "Fall"
sqrt_sub <- filter(sqrt_dat, season == seas)
log_sub <- filter(log_dat, season == seas)
```

```{r, child = '02_ANOSIM-SIMPER-NMDS.Rmd'}

```

***  
***  


# Winter  

```{r}
seas <- "Winter"
sqrt_sub <- filter(sqrt_dat, season == seas)
log_sub <- filter(log_dat, season == seas)
```

```{r, child = '02_ANOSIM-SIMPER-NMDS.Rmd'}

```

***  
***  

# Graphs of SIMPER-significant species  

These graphs are ordered from the species that show up in the most seasons to those that show up in the fewest. The subtitle for each graph names the seasons in which the species was significant.  

Each point on the graphs represents the count in a single sample event.  

Site 1 has been excluded because it was excluded from all other analyses.  

```{r}
# read in and bind all the individual data frames
spp_spr <- read.csv(here::here("output", "SIMPER_spps_Spring.csv"))
spp_sum <- read.csv(here::here("output", "SIMPER_spps_Summer.csv"))
spp_fall <- read.csv(here::here("output", "SIMPER_spps_Fall.csv"))
spp_win <- read.csv(here::here("output", "SIMPER_spps_Winter.csv"))

spp_all <- bind_rows(spp_spr, spp_sum, spp_fall, spp_win)


# group it
spp2 <- spp_all %>% 
    group_by(species) %>% 
    summarize(n_seasons = n(),
              seasons = str_c(unique(season), collapse = ", ")) %>% 
    arrange(desc(n_seasons))

# clean up counts data frame: 
# change 0s to 0.1 for log scale purposes
# turn season into a factor with a better order
# shorten the habitat type that starts with depositional
# get rid of site 1 since it wasn't in the other analyses
nfm2 <- nfm_all %>% 
    filter(site != 1) %>% 
    mutate(across(`Abudefduf.saxatilis`:ncol(nfm_all),
                  ~ case_when(. == 0 ~ 0.1,
                              . != 0 ~ as.numeric(.))),
           season = fct_relevel(season, "Spring", "Summer", "Fall", "Winter"),
           habitat_type = case_when(habitat_type == "depositional spartina edge" ~ "depositional edge",
                                    TRUE ~ habitat_type))
```


```{r, fig.height = 6, fig.width = 12}
# for each of the species in spp2..... 
# make 2 plots:
# panel for season, x-axis is habitat
# ggdist for y-axis; want all the quantities
# title = species name
# subtitle = species in which it was important in SIMPER

# one graph, raw y-axis
# the other graph, log10 y-axis


for(i in seq_along(spp2$species)){
    # species and season info
    sp_name <- spp2$species[i]
    sp_seasons <- spp2$seasons[i]
    
    # subset data frame
    sp_dat <- nfm2[c("habitat_type", "season", sp_name)]
    
    # rename for easier graphing
    names(sp_dat)[3] <- "count"
    
    # graph 1: raw y-scale
    p1 <- ggplot(sp_dat) +
        geom_dotsinterval(aes(x = habitat_type, y = count, 
                              fill = habitat_type), 
                          layout = "swarm",
                          side = "both",
                          binwidth = unit(0.05, "npc"),
                          na.rm = TRUE) +
        facet_wrap(~season, ncol = 2) +
        # scale_y_log10() + 
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              legend.position = "none") +
        labs(title = sp_name,
             subtitle = glue("SIMPER for {sp_seasons}"),
             y = "number caught",
             x = NULL)
    
    # graph 2: log10 y-scale
    p2 <- ggplot(sp_dat) +
        geom_dotsinterval(aes(x = habitat_type, y = count, 
                              fill = habitat_type), 
                          layout = "swarm",
                          side = "both",
                          binwidth = unit(0.05, "npc"),
                          na.rm = TRUE) +
        facet_wrap(~season, ncol = 2) +
        scale_y_log10(breaks = trans_breaks("log10", function(x) 10 ^ x),
                  labels = trans_format("log10", math_format(10^.x))) + 
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              legend.position = "none") +
        labs(title = sp_name,
             subtitle = glue("SIMPER for {sp_seasons}"),
             y = "number caught (log scale)",
             x = NULL) +
    annotation_logticks(sides = "l")
    
    p <- p1 + p2
    
    print(p)
}

```

