---
output: html_document
---

```{r}
set.seed(20220708)
```

Separate explanatory variables from species

```{r}
# species matrices
sqrt_spp <- select(sqrt_sub, Abudefduf.saxatilis:ones)
log_spp <- select(log_sub, Abudefduf.saxatilis:ones)

# explanatory will be the same for both data frames; they have the same rows
expl <- select(sqrt_sub, -(Abudefduf.saxatilis:ones))
```

------------------------------------------------------------------------

## ANOSIM

### Overall for `r seas`

```{r}
ano_sqrt <- anosim(sqrt_spp, grouping = expl$habitat_type,
               distance = "bray", permutations = how(blocks = expl$year_season,
                                                     nperm = 999))

ano_log <- anosim(log_spp, grouping = expl$habitat_type,
               distance = "bray", permutations = how(blocks = expl$year_season,
                                                     nperm = 999))


ano_unrestricted <- anosim(sqrt_spp, grouping = expl$habitat_type,
                           distance = "bray", permutations = 999)

```

Double-check to see if restricting permutations changed anything here.

```{r}
ano_sqrt
ano_unrestricted  
```

### Pairwise between habitats in `r seas`

```{r}
pairs <- t(combn(unique(expl$habitat_type), m = 2))
```

```{r}
pair <- character()
n <- numeric()
ano_R_sqrt <- numeric()
ano_p_sqrt <- numeric()
ano_R_log <- numeric()
ano_p_log <- numeric()
densplots <- list()

for(i in 1:nrow(pairs)){
    # which rows are in this pair  
    indcs <- which(expl$habitat_type %in% pairs[i, ])
    
    # subset data frames
    sqrt_spp_pr <- sqrt_spp[indcs, ]
    log_spp_pr <- log_spp[indcs, ]
    expl_pr <- expl[indcs, ]
    
    # run anosims  
    ano_pr_sqrt <- anosim(sqrt_spp_pr, grouping = expl_pr$habitat_type,
                          distance = "bray", permutations = how(blocks = expl_pr$year_season,
                                                                nperm = 999))

    ano_pr_log <- anosim(log_spp_pr, grouping = expl_pr$habitat_type,
                          distance = "bray", permutations = how(blocks = expl_pr$year_season,
                                                                nperm = 999))
    
    # store outputs
    pair[i] <- paste(pairs[i, ], collapse = ", ")
    n[i] <- nrow(expl_pr)
    ano_R_sqrt[i] <- ano_pr_sqrt$statistic
    ano_p_sqrt[i] <- ano_pr_sqrt$signif
    ano_R_log[i] <- ano_pr_log$statistic
    ano_p_log[i] <- ano_pr_log$signif
    
    # print the density plot of permutations  
    perm <- permustats(ano_pr_sqrt)
    samp <- glue("Sample: R = ", {round(ano_pr_sqrt$statistic, 3)}, "; p = ",
             {round(ano_pr_sqrt$signif, 4)})
    p <- densityplot(perm, main = pair[i],
            xlab = "R (permuted values)",
            ylab = "density",
            strip = FALSE,
            sub = samp)
    densplots[[i]] <- p
   
}
```

```{r}
outs <- data.frame(pair, ano_R_sqrt, ano_p_sqrt, ano_R_log, ano_p_log, n) %>% 
    arrange(desc(ano_R_sqrt))
outs$season <- seas

overall <- data.frame(
    pair = "OVERALL",
    ano_R_sqrt = ano_sqrt$statistic,
    ano_p_sqrt = ano_sqrt$signif,
    ano_R_log = ano_log$statistic,
    ano_p_log = ano_log$signif,
    n = nrow(expl),
    season = seas
)

all <- bind_rows(overall, outs)

knitr::kable(all, digits = 4)

# save data frame to csv or .rds??

```

### Density plots of permutations

```{r}
samp <- glue("Sample: R = ", {round(ano_sqrt$statistic, 3)}, "; p = ",
             {round(ano_sqrt$signif, 4)})
perm <- permustats(ano_sqrt)
densityplot(perm, 
            main = paste("OVERALL,", seas),
            xlab = "R (permuted values)",
            ylab = "density",
            strip = FALSE,
            sub = samp)
```

```{r, fig.width = 8, fig.height = 16}
# print density plots
print(densplots[[1]], split = c(1, 1, 2, 5), more = TRUE)
print(densplots[[2]], split = c(2, 1, 2, 5), more = TRUE)
print(densplots[[3]], split = c(1, 2, 2, 5), more = TRUE)
print(densplots[[4]], split = c(2, 2, 2, 5), more = TRUE)
print(densplots[[5]], split = c(1, 3, 2, 5), more = TRUE)
print(densplots[[6]], split = c(2, 3, 2, 5), more = TRUE)
print(densplots[[7]], split = c(1, 4, 2, 5), more = TRUE)
print(densplots[[8]], split = c(2, 4, 2, 5), more = TRUE)
print(densplots[[9]], split = c(1, 5, 2, 5), more = TRUE)
print(densplots[[10]], split = c(2, 5, 2, 5))
```

------------------------------------------------------------------------

## SIMPER

**Important Note**: SIMPER should only be used/interpreted if the pairwise ANOSIM for the pair of habitats was significant. Results are automatically generated for every pairwise comparison, but **not all should be used. Check ANOSIM results.**

```{r}
################################################
# SIMPER on habitat differences per season     #
# only meaningful to interpret if              #
# significant differences in pairwise ANOSIMs  #
################################################

# default simper print method prints species up to first one that drives cumulative sum above 0.7
# https://rdrr.io/cran/vegan/src/R/simper.R
# this is modified to include *all* columns from the output


# for summarizing:
top <- 4
toprint <- function(x){
  out <- character(length = top)
  for(i in 1:top){
    out[i] <- paste0(rownames(x[i, ]), " (", round(x[i, 1], 3), ")")
  }
  paste(out, collapse = "; ")
}


# SIMPER
# print top 4 species for each pairwise contrast (using function above)
# and their contribution
sqrt_simp <- simper(sqrt_spp, expl$habitat_type, 
                    permutations = how(blocks = expl$year_season,
                                                     nperm = 999))
lapply(summary(sqrt_simp), function(x) toprint(x))

# print info for all of the species that get 
# total variance explained above 70%
sqrt_simp_spp <- lapply(summary(sqrt_simp), function(z) z[seq_len(min(which(z$cumsum >= 0.7))), ])

lapply(sqrt_simp_spp, function(z) knitr::kable(z, digits = 4))
```

------------------------------------------------------------------------

## Indicator Species Analysis

According to [this online tutorial](https://jkzorz.github.io/2019/07/02/Indicator-species-analysis.html), analyses like SIMPER and indicator species analysis should give similar answers. I am trying ISA here as a complement because the SIMPER results have been difficult to interpret. BE CAUTIOUS; THIS IS FAIRLY EXPLORATORY and meant to help understand the results from ANOSIM and ordination.  

```{r}
inv = multipatt(sqrt_spp, expl$habitat_type, 
                control = how(blocks = expl$year_season,
                              nperm = 999))

summary(inv)
```


------------------------------------------------------------------------

## NMDS

#### Scree Plot

How many axes should be used?

```{r}
# from goeveg package
dimcheckMDS(sqrt_spp, k = 6, autotransform = FALSE)
```

#### run with 2 axes

```{r}
mds2 <- metaMDS(sqrt_spp, distance = "bray", k = 2,
                trymax = 50,
                autotransform = FALSE,
                trace = 0)

# print the summary info
mds2

# construct stuff for later
str2 <- glue("Stress: {round(mds2$stress, 3)}")
mds2_pts <- data.frame(mds2$points) %>% 
    bind_cols(expl)
mds2_spps <- data.frame(mds2$species)
```

#### run with 3 axes

```{r}
mds3 <- metaMDS(sqrt_spp, distance = "bray", k = 3,
                trymax = 50,
                autotransform = FALSE,
                trace = 0)

# print the summary info
mds3

# generate stuff for later plots
str3 <- glue("Stress: {round(mds3$stress, 3)}")
mds3_pts <- data.frame(mds3$points) %>% 
    bind_cols(expl)
mds3_spps <- data.frame(mds3$species)
```

#### Shepard plots and stress values

```{r}
stressplot(mds2, main = "2 axes", 
           sub = str2)
stressplot(mds3, main = "3 axes",
           sub = str3)
```

## NMDS plots

Because all the seasons (in my glance) really did need 3 dimensions in the NMDS, I'm commenting out the plots for 2-d results.

#### Points only; colored by habitat and shaped by year

```{r}
yr_shapes <- c(1, 3, 4, 6, 7, 8, 9, 10, 12, 14)
```

```{r}
# NMDS with only 2 axes


# ggplot(mds2_pts) +
#     geom_point(aes(x = MDS1, y = MDS2, color = habitat_type),
#                size = 2, alpha = 0.7) +
#     theme_bw() +
#     labs(title = glue("{seas}, 2 dimensions"),
#          subtitle = str2,
#          color = "Habitat")
# 
# ggplot(mds2_pts) +
#     geom_point(aes(x = MDS1, y = MDS2, color = habitat_type,
#                    shape = as.factor(year_sampled)),
#                size = 3) +
#     scale_shape_manual(values = yr_shapes) +
#     theme_bw() +
#     labs(title = glue("{seas}, 2 dimensions"),
#          subtitle = str2,
#          color = "Habitat",
#          shape = "Sample Year")
```

```{r}
# NMDS with 3 axes; using year as shape
# could be informative but it's a lot more to try to digest


# ggplot(mds3_pts) +
#     geom_point(aes(x = MDS1, y = MDS2, color = habitat_type),
#                size = 2, alpha = 0.7) +
#     theme_bw() +
#     labs(title = glue("{seas}, 3 dimensions"),
#          subtitle = str3,
#          color = "Habitat",
#          shape = "Sample Year")
# 
# 
# ggplot(mds3_pts) +
#     geom_point(aes(x = MDS1, y = MDS2, color = habitat_type,
#                    shape = as.factor(year_sampled)),
#                size = 3) +
#     scale_shape_manual(values = yr_shapes) +
#     theme_bw() +
#     labs(title = glue("{seas}, 3 dimensions"),
#          subtitle = str3,
#          color = "Habitat",
#          shape = "Sample Year")
```

#### Include species; without ellipses

```{r}
sqrt_simp_spp <- purrr::map(sqrt_simp_spp, rownames_to_column, var = "species") 
simps <- bind_rows(sqrt_simp_spp)
spps <- unique(simps$species)
```

```{r}
# write this out so we can make other graphs later
spps_df <- data.frame(species = spps,
                      season = rep(seas, length(spps)))
spps_out_name <- glue("SIMPER_spps_{seas}.csv")
write.csv(spps_df, here::here("output", spps_out_name), row.names = FALSE)
```

Look at all combinations of the 3 axes:

```{r}
ggord(mds3, expl$habitat_type, 
      axes = c("1", "2"), ellipse = FALSE, 
      max.overlaps = 20, var_sub = c(spps),
      repel = TRUE) +
    labs(title = glue("{seas}, axes 1 & 2"),
         subtitle = glue("3-Dimensional {str3}"),
         color = "Habitat")

ggord(mds3, expl$habitat_type, 
      axes = c("2", "3"), ellipse = FALSE, 
      max.overlaps = 20, var_sub = c(spps),
      repel = TRUE) +
    labs(title = glue("{seas}, axes 2 & 3"),
         subtitle = glue("3-Dimensional {str3}"),
         color = "Habitat")

ggord(mds3, expl$habitat_type, 
      axes = c("1", "3"), ellipse = FALSE, 
      max.overlaps = 20, var_sub = c(spps),
      repel = TRUE) +
    labs(title = glue("{seas}, axes 1 & 3"),
         subtitle = glue("3-Dimensional {str3}"),
         color = "Habitat")
```

#### Include species; with ellipses

```{r, eval = FALSE}
ggord(mds3, expl$habitat_type, axes = c("1", "2"), poly = FALSE, max.overlaps = 20, var_sub = c(spps))

ggord(mds3, expl$habitat_type, axes = c("2", "3"), poly = FALSE, max.overlaps = 20, var_sub = c(spps))

ggord(mds3, expl$habitat_type, axes = c("1", "3"), poly = FALSE, max.overlaps = 20, var_sub = c(spps))
```

## 3d NMDS plot

Set up color palette and SIMPER-only species

```{r}
# a few options I like for the palette
# viridis::viridis(n = 5, option = "C")
# viridis::viridis(n = 5, option = "H")

# RColorBrewer options aren't colorblind safe; unsure about viridis options
# RColorBrewer::brewer.pal(5, "Accent")
# RColorBrewer::brewer.pal(5, "Set1")

# see khroma package for all colorblind-friendly palettes
# as.character(khroma::colour("bright")(5))
# as.character(khroma::colour("okabeito")(5))
# as.character(khroma::colour("muted")(5))


pal <- data.frame("habitat_type" = unique(expl$habitat_type),
                  "color" = as.character(khroma::colour("bright")(5)))

# add colors to data frame
mds3_pts2 <- left_join(mds3_pts, pal)


# subset species output to only those from SIMPER
mds3_spps2 <- mds3_spps %>% 
    rownames_to_column("species") %>% 
    filter(species %in% spps)
```

```{r}

# set up points and data frames

x_pts <- mds3_pts2$MDS1
y_pts <- mds3_pts2$MDS2
z_pts <- mds3_pts2$MDS3

x_spp <- mds3_spps2$MDS1
y_spp <- mds3_spps2$MDS2
z_spp <- mds3_spps2$MDS3


### to get line segments; this is so hacky
# add a row of 0s before every row of the spps data frame
mds3_spps3 <- mds3_spps2

# work from the number of rows back to the first
mds_rows <- rev(seq(1:nrow(mds3_spps3)))

for(i in seq_along(mds_rows)){
    mds3_spps3 <- mds3_spps3 %>% 
        add_row(species = "origin", 
                MDS1 = 0, MDS2 = 0, MDS3 = 0, 
                .before = mds_rows[i])
}
```

```{r, webgl = TRUE}
# make and save the plot

# open and create size
# open3d()
par3d(windowRect = c(20, 20, 750, 750))

# plot the points
plot3d(
    x = x_pts,
    y = y_pts,
    z = z_pts,
    col = mds3_pts2$color,
    type = "s",
    size = 1,
    xlab = "MDS1",
    ylab = "MDS2",
    zlab = "MDS3"
)

# add a grid
grid3d(c("x", "y+", "z"))

# add points for the species
plot3d(
    x = x_spp,
    y = y_spp,
    z = z_spp,
    type = "s",
    size = 0.7,
    col = "black",
    add = TRUE
)

# add labels for the species
text3d(
    x = x_spp,
    y = y_spp,
    z = z_spp,
    texts = mds3_spps2$species,
    add = TRUE,
    cex = 0.8
)

# add line segments for the species
segments3d(mds3_spps3[2:4])

# add a legend 
# note these are drawn as bitmaps and can look wonky
legend3d("right", legend = pal$habitat_type, 
         pch = 16, col = pal$color, ncol = 1, 
         cex = 0.8, inset = c(0.02), text.width = NA)


# save it
out_nm <- glue("NMDS_3d_{seas}.html")
out_path <- here::here("output", out_nm)

htmlwidgets::saveWidget(rglwidget(width = 700,
                                  height = 700), 
                        file = out_path)

# close3d()
```
